---
title: "HW week 6 DOF"
author: "Angel Gallardo"
date: "2024-03-03"
output: html_document


#Abstract: Home sales in NYC for since 2010. Which neighborhood in Queens have sold the most homes since 2010? Which neghborhood have the most expensive homes, on average? 
---
```{r}
library(tidyverse)
library(dplyr)
```



```{r}
DOF <- read.csv("https://raw.githubusercontent.com/Angelogallardo05/Week6/main/DOF__Summary_of_Neighborhood_Sales_by_Neighborhood_Citywide_by_Borough_20240302.csv")
head(DOF)
```

# adjust the "TYPE.OF.HOME" in uniform 
```{r}
str(DOF)

DOF <- DOF %>%
  mutate(TYPE.OF.HOME = gsub("\\D", "", TYPE.OF.HOME))

DOF <- DOF %>%
  mutate(TYPE.OF.HOME = paste0(TYPE.OF.HOME, " Family homes"))


```



# Replace the integers in the BOROUGH column with the corroect borough name
```{r}
borough_mapping <- c("1" = "MANHATTAN", "2" = "BRONX", "3" = "BROOKLYN", "4" = "QUEENS", "5" = "STATEN ISLAND")

# Replace integers in the BOROUGH column with correct borough names
DOF$BOROUGH <- ifelse(DOF$BOROUGH %in% names(borough_mapping), 
                      borough_mapping[as.character(DOF$BOROUGH)], 
                      DOF$BOROUGH)



```

#Filter by 1 family homes in Queens

```{r}
filtered_data <- DOF %>%
  filter(TYPE.OF.HOME== "01 Family homes" & BOROUGH == "QUEENS")


```

#Bar chart to show the neighborhoods in queens with the most sales 2010-2022, one family home
```{r}
# Calculate total number of sales for each neighborhood
neighborhood_sales <- filtered_data %>%
  group_by(NEIGHBORHOOD) %>%
  summarise(total_sales = sum(NUMBER.OF.SALES)) %>%
  arrange(desc(total_sales)) %>%
  slice_head(n = 5)

# Create a bar chart
ggplot(neighborhood_sales, aes(x = reorder(NEIGHBORHOOD, -total_sales), y = total_sales)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  theme_minimal() +
  labs(x = "Neighborhood", y = "Total Number of Sales", title = "Top 5 Neighborhoods with the Most Number of Sales 20") +
  coord_flip()  # Horizontal bars for better readability
```



# 2021 included a big jumb in home sales in the North-Flushing Area, however house price did not increas dramatically.
```{r}
neighborhood_sales <- filtered_data %>%
  filter(NEIGHBORHOOD == "FLUSHING-NORTH") %>%
  group_by(YEAR)
print(neighborhood_sales)

plot_sales <- ggplot(neighborhood_sales, aes(x = YEAR, y = NUMBER.OF.SALES)) +
  geom_line(color = "blue") +
  labs(x = "Year", y = "Total Number of Sales", title = "Total Number of Sales per Year for FLUSHING-NORTH") +
  theme_minimal()

# Create plot for Average Sale Price
plot_price <- ggplot(neighborhood_sales, aes(x = YEAR, y = AVERAGE.SALE.PRICE/1000000)) +
  geom_line(color = "red") +
  labs(x = "Year", y = "Average Sale Price (MM)", title = "Average Sale Price per Year for FLUSHING-NORTH 2010-2022") +
  theme_minimal()

# Display both plots
plot_sales
plot_price
```



# LIC had a large decline in average home sale price price starting in 2021
```{r}
top5_data <- filtered_data %>%
  group_by(NEIGHBORHOOD) %>%
  summarise(avg_sale_price = mean(AVERAGE.SALE.PRICE)) %>%
  top_n(5, avg_sale_price) %>%
  inner_join(filtered_data, by = "NEIGHBORHOOD")  # Join with original data to get all years



# Create a line plot for top 5 neighborhoods
ggplot(top5_data, aes(x = YEAR, y = AVERAGE.SALE.PRICE, color = NEIGHBORHOOD)) +
  geom_line() +
  scale_x_continuous(breaks = unique(top5_data$YEAR), labels = unique(top5_data$YEAR)) +
   scale_y_continuous(labels = scales::dollar_format(suffix = "MM", scale = 1e-6))+
  theme_minimal() +
  labs(x = "Year", y = "Average Sale Price", title = "Average Single Home Sale Price by Year for Top 5 Neighborhoods") +
  theme(legend.position = "right")
```

