---
title: "HW6 1of3 DATA607"
author: "Angel Gallardo"
date: "2024-03-02"
output: html_document
---

#ANgel Gallardo 

# analyze nyc monthly jobs by industry for the last 34 years. Which industry provide more stability? WHat is the growth by industry for the last 10 years?


```{r}
library(tidyverse)
library(dplyr)
```


```{r}
nycemp <- read.csv("https://raw.githubusercontent.com/Angelogallardo05/Week6/main/nychist.csv")
head(nycemp)
```



# make the columns with numbers as numeric.
```{r}



nycemp <- nycemp %>%
  mutate(across(starts_with("JAN"):ends_with("ANNUAL"), ~as.numeric(gsub(",", "", .))))

head(nycemp)

```

# remove other columns

```{r}
nycemp <- nycemp %>%
  select(-19:-47)
         

```

#fill the average colu,m
```{r}
nycemp <- nycemp %>%
  rowwise() %>%
  mutate(ANNUAL = ifelse(is.na(ANNUAL), mean(c_across(JAN:DEC), na.rm = TRUE), ANNUAL)) %>%
  ungroup()

head(nycemp)
```








# pivot wider to make the Years into columns

```{r}
nycemp_year <- nycemp %>%
  group_by(INDUSTRY_TITLE, AREANAME, YEAR) %>%
  summarise(ANNUAL = mean(ANNUAL)) %>%
  select(INDUSTRY_TITLE, AREANAME, YEAR, ANNUAL)

nycemp_filtered <- nycemp_year %>%
  pivot_wider(names_from = YEAR, values_from = ANNUAL) %>%
  mutate_all(~ replace(., is.na(.), 0)) %>%
  select(INDUSTRY_TITLE, AREANAME, order(as.numeric(colnames(.))))  



```





# calculate the average job growth rate for each industry in the past 10 years.
```{r}
last_10_years <- nycemp_filtered %>%
  select(INDUSTRY_TITLE, AREANAME, `2014`:`2023`)

# Calculate percentage changes for each year
percentage_changes <- last_10_years %>%
  mutate(
    Percentage_Change_2015 = ((`2015` - `2014`) / `2014`) * 100,
    Percentage_Change_2016 = ((`2016` - `2015`) / `2015`) * 100,
    Percentage_Change_2017 = ((`2017` - `2016`) / `2016`) * 100,
    Percentage_Change_2018 = ((`2018` - `2017`) / `2017`) * 100,
    Percentage_Change_2019 = ((`2019` - `2018`) / `2018`) * 100,
    Percentage_Change_2020 = ((`2020` - `2019`) / `2019`) * 100,
    Percentage_Change_2021 = ((`2021` - `2020`) / `2020`) * 100,
    Percentage_Change_2022 = ((`2022` - `2021`) / `2021`) * 100,
    Percentage_Change_2023 = ((`2023` - `2022`) / `2022`) * 100
  ) %>%
  select(-matches("^\\d{4}$"))  # Remove original columns


```


```{r}
```

# sort by highest grwoth in 2023

```{r}
highest_growth_2023 <- percentage_changes %>%
  filter(Percentage_Change_2023 == max(Percentage_Change_2023)) %>%
  arrange(desc(Percentage_Change_2023))



glimpse(percentage_changes)
```
```{r}
percentage_changes <- percentage_changes %>%
  rename_with(~ gsub("Percentage_Change_", "", .), starts_with("Percentage_Change_"))

```




# Home Health care services have the least violatile job growth in the last 10 years

```{r}
top_5_industries <- highest_growth_2023$INDUSTRY_TITLE[1:5]
top_5_growth <- percentage_changes %>%
  filter(INDUSTRY_TITLE %in% top_5_industries) %>%
  pivot_longer(cols = starts_with("2"), 
               names_to = "Year", 
               values_to = "Percentage_Change")

# Plot
ggplot(top_5_growth, aes(x = Year, y = Percentage_Change, group = INDUSTRY_TITLE, color = INDUSTRY_TITLE)) +
  geom_line() +
  geom_point() +
  labs(title = "Top 5 Industries with Highest Growth Percentage (2014-2023)",
       x = "Year",
       y = "Percentage Change") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```





